Modern Statistical Methods --- Example Sheet 2

### 2020-11-08 inverse_problems/C3.tex
lots of work done, lots of confusion going on

\begin{theorem}
	Let $\XX, \YY$ be Banach spaces with topologies $\tau_\XX, \tau_\YY$. Assume that:
	\begin{enumerate}[(a)]
		\item bounded sequences in $\XX$ have $\tau_\XX$-convergent subsequences;
		\item $\JJ\colon \XX \to [0, \infty]$ is proper, convex, $\tau_\XX$-LSC and satisfies the assumptions of \cref{lem:regulariser_coercive};
		\item $A$ is continuous w.r.t.\ $\tau_\XX$ and $\tau_\YY$;
		\item $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC. 
	\end{enumerate}

@@

>>>>

\begin{theorem} \label{thm:minimiser_existence}
	Let $\XX, \YY$ be Banach spaces with topologies $\tau_\XX, \tau_\YY$. Assume that:
	\begin{enumerate}[(a)]
		\item bounded sequences in $\XX$ have $\tau_\XX$-convergent subsequences;
		\item $\JJ\colon \XX \to [0, \infty]$ is proper, convex, $\tau_\XX$-LSC and satisfies the assumptions of \cref{lem:regulariser_coercive};
		\item $A$ is continuous w.r.t.\ $\tau_\XX$ and $\tau_\YY$;
		\item $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC. 
	\end{enumerate}

@@

\begin{theorem}
	Let the assumptions of \cref{thm:minimiser_existence} be satisfied and suppose that $\inf\limits_{v \in \XX} \norm{Av - f} = 0$. 
	Let $\alpha = \alpha(\delta)$ satisfy \[
	\lim_{\delta\to0} \alpha(\delta) = 0  = \limsup_{\delta \to 0} \frac{\delta^2}{\alpha(\delta)}. 
	\]
	Then any sequence $(\delta_n) \to 0$ has a subsequence $(\delta_n) \to 0$ such that $u_{\delta_n} \ceq u_{\alpha(\delta_n)} \to u_\JJ \D$ and $\JJ(u_\delta) \to \JJ(u_\JJ\D)$, where $u_\JJ\D$ is a $\JJ$-minimising solution. 
\end{theorem}

\begin{proof} 
	By \cref{lem:regulariser_coercive}, we know that $\Phi_\alpha$ is coercive, and by \cref{thm:minimiser_existence}, we know there exists a $\JJ$-minimising solution $u_\JJ\D$. For any $(u_{\delta_n})$ with $\delta_n \to 0$ we find
	\begin{equation} \label{eq:phi_alpha}
	\Phi_\alpha(u_{\delta_n}) \leq \Phi_\alpha(u_\JJ\D) = \frac12\norm{A u_\JJ\D - f_\delta}^2 + \alpha(\delta) \JJ(u_\JJ\D) \leq \frac{\delta^2}{2} + \alpha(\delta)\JJ(u_\JJ\D),
	\end{equation}
	which is bounded, so by coercivity of $\Phi_\alpha$ we conclude that $(u_{\delta_n})$ is bounded.

	After taking a subsequence if necessary, we conclude that $u_{\delta_n} \overset{\tau_X}\to u_0$. We show that $u_0$ is a $\JJ$-minimising solution. Since $A \colon \tau_X \to \tau_Y$ is continuous and $\norm{\cdot}_\YY$ is $\tau_\YY$-LSC, we get
	\[
	\norm{Au_0 - f} \leq \liminf_{n\to\infty} \norm{Au_{\delta_n} - f} \leq \liminf_{n\to\infty} \norm{Au_{\delta_n} - f_{\delta_n}} + \norm{f - f_{\delta_n}} = 0,
	\]
	since $\norm{Au_{\delta_n} - f\delta_n} \leq 2 \Phi_\alpha(u_{\delta_n}) \to 0$ by \cref{eq:phi_alpha}. This shows that $u_0$ is a least-squares solution.

	Furthermore, we have 
	\[
	\JJ(u_0) \leq \liminf_{n\to\infty} \JJ(u_{\delta_n}) \leq \limsup_{n\to\infty} \JJ(u_{\delta_n}) \leq \limsup_{n\to\infty} \frac{\Phi_\alpha(u_{\delta_n})}{\alpha(\delta)} \leq \limsup_{n\to\infty} \frac12 \frac{\delta_n^2}{\alpha(\delta_n)} + \JJ(u_\JJ\D) = \JJ(u_\JJ\D)
	\]
	which proves that $u_0$ minimises $\JJ$.

	To prove that $u_{\delta_n}$ is actually a minimising sequence, note that all above inequalities are in fact equalities, so $\liminf_{n\to\infty} \JJ(u_{\delta_n}) = \limsup_{n\to\infty} \JJ(u_{\delta_n}) = \lim_{n\to\infty} \JJ(u_{\delta_n}) = \JJ(u_\JJ\D)$.

%	By the previous theorem, we know there exists a $\JJ$-minimising solution $u_0$, and for every $\alpha > 0$ we know there exists
%	\[
%	u_\delta \ceq u_{\alpha(\delta)} \in \argmin_u \frac12 \norm{Au - f}^2 + \alpha \JJ(u). 
%	\]
%	
%	Therefore, we have 
%	\begin{equation} \label{eq:estimate}
%	\frac12 \norm{Au_\delta - f_\delta}^2 + \alpha(\delta) \JJ(u_\delta) \leq \frac12 \norm{A u_0 - f_\delta}^2 + \alpha(\delta) \JJ(u_0) \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0). 
%	\end{equation}
%	\Cref{eq:estimate} gives us two estimates: the first one is
%	\[
%	\alpha(\delta) \JJ(u_\delta) \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0) \implies \JJ(u_\delta) \leq \frac12 \frac{\delta^2}{\alpha(\delta)} + \JJ(u_0), 
%	\]
%	so $\JJ(u_\delta)$ is bounded uniformly in $\delta$. From \cref{eq:estimate} we also infer
%	\[
%	\frac12 \norm{Au_\delta - f_\delta}^2 \leq \frac{\delta^2}{2} + \alpha(\delta) \JJ(u_0),
%	\]
%	from which it follows that $\norm{Au_\delta - f_\delta}$ is also uniformly bounded in $\delta$. 
%	
%	Writing $u_\delta = u_\delta^0 + u_\delta^{\Cal N}$
\end{proof}

\subsection{Total variation regularisation}
\begin{definition}
	Let $\Omega \subseteq \RR^n$ be a bounded comain and $u \in L^1(\Omega)$. Let
	\[
	\DD(\Omega, \RR^n) \ceq \qty{\phi \in C_0^\infty(\Omega, \RR^n) \mid \sup_{x \in \Omega} \norm{\phi(x)}_2 \leq 1}. 
	\]
	We define the \emph{total variation} of $u \in L^1(\Omega)$ as
	\[
	\TV(u) \ceq \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u(x) (\div \phi)(x) \dd{x}. 
	\]
\end{definition}

\begin{example}
	If $u \in L^1(\Omega)$ has a weak derivative $\grad u \in L^1$ (i.e., $u \in W^{1, 1}$), then since any $\phi \in \DD(\Omega, \RR^n)$ is 0 on the boundary of $\Omega$, we have
	\[
	\TV(u) = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u(x) (\div\phi)(x) \dd{x} =  \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega -\grad u \vdot \phi \dd{x}, 
	\]
	By Cauchy-Schwarz we have $\abs{\nabla u(x) \vdot \phi(x)} \leq \norm{\nabla u(x)}_2 \norm{\phi(x)}_2$ which gives an upper bound for $\TV(u)$, and it can be shown that this upper bound is indeed an equality, so
	\[
	\TV(u) = \int_\Omega \norm{\nabla u}_2 \dd{x} = \norm{\nabla u}_{L^1}.
	\]
\end{example}

\begin{proposition}
	The functional $\TV$ is proper, convex, and absolutely one-homogeneous, and for any constant $c$ and any $u \in L^1(\Omega)$ we have
	\[
	\TV(c) = 0, \quad \TV(u + c) = \TV(u). 
	\]
\end{proposition}

\begin{proof}
	For any test function $\phi$ we have 
	\[
	\int_\Omega \div\phi \dd{x} = \int_{\pt\Omega} \phi \vdot \vec n \dd{x} = 0
	\]
	since $\phi = 0$ on the boundary. This shows immediately that $\TV(c) = 0$ and that $\TV(u + c) = \TV(u)$. 
\end{proof}

\begin{remark}
	In fact, the converse also holds: $\TV(u) = 0$ if and only if $u$ is constant. 
\end{remark}

\begin{example}
	Let $C \subseteq \Omega$ be a bounded domain with smooth boundary. Then we have
	\[
	\TV(\ind_C) = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_C\div\phi \dd{x} = \sup_{\phi \in \DD(\Omega, \RR^n)} \int_{\pt C} \phi \vdot \vec n \dd{s} \leq \sup_{\phi \in \DD(\Omega, \RR^n)} \int_{\pt C} \norm{\phi}_2 \norm{\vec n}_2  \dd{s} = \int_{\pt C} 1 \dd{s},
	\]
	and we can show that the inequality here is in fact an equality, i.e., $\TV(\ind_C) = \int_{\pt C} 1 \dd{s}$, which is the perimeter of $C$. 
\end{example}

\begin{definition}
	We define the space of functions of \emph{bounded variation} as
	\[
	\BV(\Omega) \ceq \qty{u \in L^1(\Omega) : \norm{u}_{\BV} \ceq \norm{u}_{L^1} + \TV(u) < \infty}.
	\]
\end{definition}
\begin{remark}
	It can be shown that BV is the dual of a separable Banach space, and that weak-$*$ convergence in BV is equivalent to strong convergence in $L^1$ and convergence of the values $\TV(u_n) \to \TV(u)$. 
\end{remark}

\begin{definition}
	We define a \emph{domain} in $\RR^n$ as an open connected subset, and it is called a \emph{Lipschitz domain} if its boundary is locally the graph of a Lipschitz continuous function (see Wikipedia for the details). 
\end{definition}

\begin{theorem}
	Let $\Omega \subseteq \RR^n$ be a bounded Lipschitz domain and $p, m \in \NN$. Define 
	\[
	p^* \ceq \begin{cases} \frac{np}{n - mp}, &n > mp, \\ \infty,& n \leq mp. \end{cases}
	\]
	Then the embedding $W^{m, p} \to L^q(\Omega)$ is continuous for all $q \in [1, p^*]$ and compact for all $q \in [1, p^*)$. 
\end{theorem}

A consequence ($p = m = 1$) is the following:
\begin{corollary}
	For any bounded Lipschitz domain $\Omega \subseteq \RR^n$, the $\BV(\Omega)$ is compactly embedded in $L^1(\Omega)$ for $n \geq 2$. For $n = 2$, the embedding $\BV(\Omega) \to L^2(\Omega)$ is continuous. 
\end{corollary}

\begin{theorem}
	Let $\Omega \subseteq \RR^n$ be open and bounded. Then $\TV$ is LSC in $L^1(\Omega)$. 
\end{theorem}

\begin{proof}
	Let $(u_j) \subseteq \BV(\Omega)$ converge to $u \in L^1(\Omega)$. Then for any test function we have
	\begin{align*}
		\TV(u) &= \sup_{\phi \in \DD(\Omega, \RR^n)}\int_\Omega u_j(x) (\div \phi)(x) \dd{x} \\
		&= \sup_{\phi \in \DD(\Omega, \RR^n)} \lim_{j\to\infty} \int_\Omega u_j(x) (\div \phi)(x) \dd{x} \\
		&\leq \liminf_{j\to\infty} \sup_{\phi \in \DD(\Omega, \RR^n)} \int_\Omega u_j(x) (\div\phi)(x) \dd{x} \\
		&= \liminf_{j\to\infty} \TV(u_j). 
	\end{align*}
\end{proof}

### 2020-11-08 modern_statistical_methods/C2.tex
lots of work done, lots of confusion going on

We have the following:
\begin{theorem}[Slow rate]
	Assume $X$ has centred and scaled columns, and assume that $Y$ has been centred, so $Y = X\beta^0 + \eps - \bar\eps \vec 1$. Let $A > 0$ and suppose
	\[
	\lambda = A\sigma\sqrt{\frac{\log(p)}{n}}.
	\]
	Let $\hat\beta = \hat\beta_\lambda^\Rm{L}$, then with probability at least $1 - 2p^{-(A^2/2 - 1)}$ we have that the MSPE satisfies
	\[
	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 \leq 4\lambda \norm{\beta^0}_1 = 4A\sigma\sqrt{\frac{\log(p)}{n}} \norm{\beta^0}_1. 
	\]
\end{theorem}

@@

\begin{lemma} \label{lem:probability_omega_bound}
	Suppose $\eps_1, \dotsc, \eps_n$ are independent mean-zero $\sigma$-sub-Gaussian random variables and let $\lambda \ceq A\sigma \sqrt{\log(p)/n}$. Then
	\[
	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} \leq \lambda) \geq 1 - 2p^{-(A^2/2 - 1)}. 
	\]
\end{lemma}

@@

\subsubsection{Optimisation theory and convex analysis}
\begin{proposition}
	Let $f \colon \RR^d \to \RR$ be twice continuously differentiable. Then $f$ is convex (resp.\ strictly convex) if and only if its Hessian is positive semi-definite (resp.\ positive definite) for all $x \in \RR^d$.
\end{proposition}

>>>>

We have the following bound on the prediction error:
\begin{theorem}[Slow rate] \label{thm:lasso_slow_rate}
	Assume $X$ has centred and scaled columns, and assume that $Y$ has been centred, so $Y = X\beta^0 + \eps - \bar\eps \vec 1$. Let $A > 0$ and suppose
	\[
	\lambda = A\sigma\sqrt{\frac{\log(p)}{n}}.
	\]
	Let $\hat\beta = \hat\beta_\lambda^\Rm{L}$, then with probability at least $1 - 2p^{-(A^2/2 - 1)}$ we have that the MSPE satisfies
	\[
	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 \leq 4\lambda \norm{\beta^0}_1 = 4A\sigma\sqrt{\frac{\log(p)}{n}} \norm{\beta^0}_1. 
	\]
\end{theorem}

@@

\begin{lemma} \label{lem:probability_omega_bound}
	Suppose $\eps_1, \dotsc, \eps_n$ are independent mean-zero $\sigma$-sub-Gaussian random variables and let $\lambda \ceq A\sigma \sqrt{\log(p)/n}$. Then
	\[
	\PP\qty(\frac{\norm{X\T\eps}_\infty}{n} \leq \lambda) \geq 1 - 2p\exp(\frac{-\lambda^2}{2(\sigma/\sqrt n)^2}) =  1 - 2p^{-(A^2/2 - 1)}. 
	\]
\end{lemma}

@@

\subsubsection{Optimisation theory and convex analysis}
\subsubsection*{Lagrangian optimisation}
\begin{definition}
	For the optimisation problem
	\begin{equation} \label{eq:opt_primal}
	\inf \qty{f(x) \mid g(x) = 0, h(x) \leq 0},
	\end{equation}
 	its \emph{Lagrangian} is given by
 	\[
 	L(x, \lambda,\nu) \ceq f(x) + \lambda h(x) + \nu g(x),
 	\]
 	and its \emph{Lagrange dual function} is
 	\begin{equation} \label{eq:opt_dual}
 	\tilde f(\lambda, \nu) = \inf_x L(x, \lambda, \nu). 
 	\end{equation}
\end{definition}

Note that if $p^*$ is the minimum value of \eqref{eq:opt_primal}, then for any feasible $x$ (i.e., $x$ with $g(x) = 0$, $h(x) \leq 0$)
we have for $\lambda \geq 0$ that 
\[
f(x) \geq L(x, \lambda, \nu) \geq \tilde f(\lambda, \nu).
\]
By taking the infimum over $x$ in the left-hand side and the supremum over $\lambda, \nu$ in the right-hand side we find that
\[
p^*  \geq \sup_{\lambda \geq 0, \nu} \tilde f(\lambda, \nu),
\]
Sometimes we actually have equality here, and in this case we say that the \emph{duality gap} is zero. Note that $\tilde f$ is a concave function, and is therefore easy to maximise.

A sufficient condition for a zero duality gap is that $L(\cdot, \lambda, \nu)$ is convex and that the problem is feasible.

\subsubsection*{Convex analysis}

Let $\rext = \RR \cup \qty{\infty}$, and define for any $f \colon \RR^d \to \rext$ its (\emph{effective}) \emph{domain} $\dom(f) \ceq \qty{x \mid f(x) < \infty}$.
\begin{proposition}
	Let $f \colon \RR^d \to \RR$ be twice continuously differentiable. Then $f$ is convex (resp.\ strictly convex) if and only if its Hessian is positive semi-definite (resp.\ positive definite) for all $x \in \RR^d$.
\end{proposition}

\begin{definition}
	Let $f \colon \RR^d \to \rext$  be convex and $x \in \RR^d$. Then $v \in \RR^d$ is called a \emph{subgradient} of $f$ at $x$ if
	\[
	f(y) \geq f(x) + v\T(y-x) \qquad\forall y \in \RR^d. 
	\]
	The set of all subgradients of $f$ at $x$ is called the \emph{subdifferential} of $f$ at $x$, denoted by $\pt f(x)$. 
\end{definition}

\begin{proposition}
	If $f$ is convex and differentiable at $x \in \Int(\dom f)$, then $\pt f(x) = \qty{\nabla f(x)}$. 
\end{proposition}

\begin{proposition}
	Let $f, g \colon \RR^d \to \rext$ be convex with $\Int(\dom f) \cap \Int(\dom g) \neq\emptyset$, $x \in \RR^d$ and $\alpha > 0$. Then $\pt(\alpha f)(x) = \alpha \pt f(x)$ and $\pt(f+g)(x) = \pt f(x) + \pt g(x)$. 
\end{proposition}

\begin{proposition}[KKT conditions]
	Let $f \colon \RR^d \to \rext$ be convex. Then $x^*$ is a global minimiser of $f$ if and only if $0 \in \pt f(x^*)$. 
\end{proposition}

\begin{proof}
	Clearly $x^*$ is a global minimiser of $f$ if and only if $f(y) \geq f(x^*) + 0\T (y - x)$ for all $y \in \RR^d$. 
\end{proof}

We define for $x \in \RR$
\[
\sign(x) \ceq\begin{cases}
	-1, &x<0,\\ 0, &x=0, \\ 1, &x>0,
\end{cases}
\]
and for $x \in \RR^d$ we define $\sign(x) \ceq (\sign(x_1), \dotsc, \sign(x_n))$.

\begin{proposition}
	Let $x \in \RR^d$ and $A \ceq \qty{j \mid x_j \neq 0}$. Then
	\[
	\pt \norm{x}_1 = \qty{v \in \RR^d : \norm{v}_\infty \leq 1 \text{ and } v_A = \sign(x_A)}. 
	\]
\end{proposition}

\begin{proof}
	Define $g_j(x) = \abs{x_j}$ so that $\norm{\cdot}_1 = \sum_j g_j$ and $\pt \norm{x}_1 = \sum_j \pt g_j(x)$. Now, if $x_j \neq 0$, then $g_j$ is differentiable at $x$ so $\pt g_j(x) = \qty{\sign(x_j) e_j}$. If $x_j = 0$, then it can be shown that
	\[
	\pt g_j(x) = \qty{c e_j \mid c \in [-1, 1]}.
	\]
	Combining the above facts proves the claim. 
\end{proof}

With the previous proposition, we can write characterise solutions to the Lasso: namely we have
\[
\hat\beta_\lambda^L = \argmin_\beta \frac1{2n} \norm{Y- X\beta}_2^2 + \lambda \norm{\beta}_1 = \argmin_\beta Q(\lambda)
\]
if and only if $0 \in \pt Q(\hat \beta_\lambda^L)$. A simple computation gives, with $A = \qty{k \mid \beta_k \neq 0}$, 
\begin{align*}
\pt Q(\beta) &= \pt(\frac1{2n} \norm{Y- X\beta}_2^2) + \pt(\lambda \norm{\beta}_1) \\
&= \qty{-\frac1n X\T(Y - X\beta)} + \qty{\lambda \nu : \norm{\nu}_\infty \leq 1, \nu_A = \sign(\beta_A)}. 
\end{align*}

We conclude that $\hat\beta$ is a Lasso solution if and only if there exists a vector $\hat\nu$ with $\norm{\hat\nu}_{\infty} \leq 1$ and $\hat\nu_A = \sign(\hat\beta_A)$ such that
\[
\frac1n X\T(Y - X\hat\beta) = \lambda \hat\nu. 
\]

Note that $\hat\beta^L_\lambda$ need not be unique, however, the fitted values are unique:
\begin{proposition}
	$X\hat\beta_\lambda^L$ is unique. 
\end{proposition}

\begin{proof}
	Write $\frac1{2n}\norm{Y- X\beta}_2^2 = Q_1(X \beta)$ and $\norm{\beta}_1 = Q_2(\beta)$. Note that $Q_1$ is strictly convex. Now, for any $t\in (0, 1)$ and any two lasso solutions $\hat\beta, \hat\gamma$ with $Q(\hat\beta) = Q(\hat\gamma) = c^*$ we have

	\begin{align*}
		c^* &\leq Q(t\hat\beta + (1-t)\hat\gamma) = Q_1(tX\hat\beta+(1-t)X\hat\gamma) + Q_2(t\hat\beta + (1-t)\hat\gamma) \\
		&\overset\star\leq t Q_1(X\hat\beta) + (1-t)Q_2(X\hat\gamma) + Q_2(t\hat\beta + (1-t)\hat\gamma) \\
		&\leq t Q(\hat\beta) + (1-t)Q(\hat\gamma) = c^*,
	\end{align*}
so all inequalities must be equalities. In particular, by $\star$ we find $Q_1(tX\hat\beta + (1-t)X\hat\gamma) = t Q_1(X\hat\beta) + (1-t)Q_1(X\hat\gamma)$, and since $Q_1$ is strictly convex this implies $X\hat\beta = X\hat\gamma$. 
\end{proof}

\subsection{Variable selection}
We consider the noiseless model $Y = X\beta^0$, where no randomness is involved. Let $S = \qty{k \mid \beta_k \neq 0}$, $N = \qty{1, \dotsc, p} \setminus S$, and assume that $S = \qty{1, \dotsc, s}$ (so $\beta_1, \dotsc, \beta_s$ are nonzero and $\beta_{s+1}, \dotsc, \beta_p$ are zero).

Furthermore, write $X = \mqty[ X_S & X_N]$ and assume that $\rank(X_S) = s$, so that $X_S\T X_S$ is invertible (NB in particular this implies $n \geq s$). 
\begin{theorem}
Let $\lambda > 0$ and define $\Delta = X_N\T X_S(X_S\T X_S)^{-1} \sign(\beta^0_S)$. 
\begin{enumerate}
	\item If $\norm{\Delta}_{\infty} \leq 1$ and for $k = 1, \dotsc, s$ we have
	\[
	\abs{\beta^0_k} > \lambda \abs{\sign(\beta^0_S)\T \qty[\frac1n X_S\T X_S]^{-1}_k},
	\]
	then there exists a Lasso solution $\hat\beta_\lambda^L$ with $\sign(\hat\beta_\lambda^L) = \sign(\beta^0)$.

	\item If there exists a Lasso solution $\hat\beta_\lambda^L$ with $\sign(\hat\beta_\lambda^L) = \sign(\beta^0)$, then $\norm{\Delta}_{\infty} = 1$. 
\end{enumerate}
\end{theorem}

\begin{remark}
	The condition $\norm{\Delta}_{\infty} \leq 1$ is called the \emph{irrepresentable condition}, and it roughly states that none of the insignificant predictors align ``too well'' with the response. \TODO Elaborate on this (lecture 12). 
\end{remark}
\begin{proof}
	Let $\hat\beta$ be any lasso solution and write $\hat S = \qty{k \mid \beta_k \neq 0}$. Using the fact that $\beta^0_N = 0$, the KKT conditions for the Lasso can be expanded as
	\begin{equation} \label{eq:expanded_lasso_conditions}
	\frac1n \mqty(X_S\T X_S & X_S\T X_N \\ X_N\T X_S & X_N\T X_N) \mqty(\beta^0_S - \hat\beta_S\\ -\hat\beta_N) = \lambda \mqty(\hat\nu_S\\ \hat\nu_N)
	\end{equation}
	for some $\hat\nu$ with $\norm{\hat\nu}_\infty \leq 1$ and $\hat\nu_{\hat S} = \sign(\hat\beta_{\hat S})$. 
	\begin{enumerate}
		\item It is easily checked that taking
		\begin{align*}
		(\hat\beta_S, \hat\beta_N) &= \qty(\beta^0_S - \lambda \qty(\frac1n X_S\T X_S)^{-1} \sign(\beta^0_S), 0) \\
		(\hat\nu_S, \hat\nu_N) &= (\sign(\beta^0_S), \Delta)
		\end{align*}
		is a Lasso solution with $\sign(\hat\beta^0_S) = \sign(\hat\beta_S)$. 
		\item If $\sign(\hat\beta) = \sign(\beta^0)$, then $\hat S = S$, $\hat \nu_S= \sign(\beta^0_S)$ and $\hat\beta_N = 0$. The top block of \cref{eq:expanded_lasso_conditions} can be written as
		\[
		\frac1n X_S\T X_S (\beta^0_S - \hat\beta_S) = \lambda \sign(\beta^0_S) \implies \beta^0_S - \hat\beta_S = \lambda \qty[\frac1n X_S\T X_S]^{-1} \sign(\beta^0_S). 
		\]
		Plugging that into the bottom block of the same equation, we get
		\[
		\lambda \frac1n X_N\T X_S \qty[\frac1nX_S\T X_S]^{-1} \sign(\beta^0_S) = \lambda \hat\nu_N \implies \norm{\Delta}_\infty \leq 1. 
		\]
	\end{enumerate}
\end{proof}

\subsubsection{Prediction and estimation}
\begin{definition}
	Let $X \in \RR^{n \times p}$ be a design matrix and $S \subseteq \qty{1, \dotsc, p}$ where $s \ceq \abs{S} > 0$, then the \emph{compatibility constant} $\phi^2 = \phi^2(X, S)$ is defined as
	\[
	\phi^2 \ceq \inf\qty{\frac{\frac1n \norm{X\beta}_2^2}{\frac1s \norm{\beta_S}_1^2} : \beta \in \RR^p, \beta_S \neq 0, \norm{\beta_N}_1 \leq 3 \norm{\beta_S}_1}. 
	\]
	The \emph{compatibility condition} is that $\phi^2 > 0$. 
\end{definition}

Note that if $X_S$ is rank-deficient, then there exists  $\beta_S \neq 0$ with $X_S\beta_S = 0$, and by setting $\beta_N = 0$ we find that $\phi^2 = 0$. Therefore, $X_S$ must have full column rank for the compatibility condition to be satisfied, and in particular we must have $n \geq s$.

\begin{proposition}
	If $X\T X/n$ has minimal eigenvalue $c_\Rm{min}$ then $\phi^2 \geq c_\Rm{min}$. In particular, if $X$ has full column rank (i.e., $c_\Rm{min} > 0$), the compatibility condition is always satisfied. 
\end{proposition}

\begin{proof}
	We have
	\[
	\norm{\beta_S}_1 = \sign(\beta_S)\T \beta_S\leq \norm{\sign(\beta_S)}_2 \norm{\beta_S}_2 = \sqrt{s} \norm{\beta}_2.
	\]
	Therefore, we find
	\[
	\phi^2 \geq \inf_{\beta \neq 0}  \frac{\frac1n \norm{X\beta}_2^2}{\frac1s \norm{\beta}_1^2} \geq \frac1n \inf_{\beta \neq 0} \frac{\norm{X\beta}_2^2}{\norm{\beta}_2^2} = \inf_{\norm{\beta} =1 } \beta\T \qty(\frac{X\T X}{n}) \beta = c_\Rm{min}. 
	\]
\end{proof}

Note that the converse of this proposition does not necessarily hold: $X$ does not need to have full column rank to satisfy the compatibility condition. In particular, it does not need to be the case that $n < p$.

Now we present the ``fast'' convergence rate of the lasso (compare with \cref{thm:lasso_slow_rate}):
\begin{theorem}[Fast rate] \label{thm:lasso_fast_rate}
	Let the conditions of \cref{thm:lasso_slow_rate} be satisfied and assume furthermore that the compatibility condition holds. 
\end{theorem}
